\Large \bf Motivating a Constraint-Based Approach to Colliding Hash Functions

### 2018-02-02 honors-thesis/usenix.tex

\section{Introduction}
\subsection{History of SAT in Cryptography}

@@

A collision is \textbf{extensible} if one example of a collision will lead to
other examples under the same differential and weak state transition constraints.
Note that an extensible collision is inherently constraint-based.

A collision is \textbf{differential-independent} if there are multiple
differentials which satisfy the intermediate state constr
A collision is \textbf{malleable} if collisions can be found with different
properties such as ASCII or JSON input blocks, or with fixed internal text.

A collision is \textbf{state-independent} if there exist alternative
starting states which produce collisions. A state-independent collision
is termed strong if one collision can collide under two or more starting
states.

@@

We now introduce a series of notations to ease construction of cryptographic
SAT models. $b := AND_{i = 0}^{32}(a)$. Needs to be

@@

\subsubsection{Constraint-Based}
\begin{itemize}
    \item Simple encoding
    \item Advanced constraints
    \item Time difference
\end{itemize}
testing
\subsubsection{Differential-Independent}
\subsubsection{State-Independent}
\subsubsection{Malleable}
\begin{itemize}
    \item ASCII
    \item JSON
    \item substring
\end{itemize}
Testing
\subsubsection{Additional Notes}

@@

\section{MD5}
\subsection{Brief History and Overview}
\subsection{Wang's MD5 Collisions}
\subsection{Stevens's MD5 Collisions}
\subsection{Xie's MD5 Collisions}

>>>>

\section{Introduction}
\subsection{Brief History of SAT in Cryptography}

Much of the original research for developing the use of SAT with cryptography
was done by Fabio Massacci currently at the University of Trento \cite{CryptoSAT}.
The focus of this work was largely two fold: cryptoanalysis against the DES
and faking signatures in the RSA with $e=3$. More information can be found
here \cite{MASSACCI2000LogicalCA} and here \cite{FIORINI2003101}, but note that
this work predates all of the collisions discussed in this article.

@@

A collision is \textbf{extensible} if one example of a collision will lead to
other examples under the same differential and weak state transition constraints.
Note that if a collision is extensible, it is also constraint-based.

A collision is \textbf{differential-independent} if there are multiple
differentials which satisfy the state transition constraints.

A collision is \textbf{malleable} if collisions can be found with different
properties such as ASCII or JSON input blocks, or with fixed internal text.

A collision is \textbf{state-independent} if there exist alternative
starting states which produce collisions. A state-independent collision
is termed \textbf{strong} if one block can collide under two or more starting
states.

@@

We now introduce a series of notations to ease construction of cryptographic
SAT models.

We define
\begin{align*}
    c_{l}^{l + k - j} := EQ_{i = j}^{k}(a_i, b_i)
\end{align*}
for $j \geq 0, k \geq j, l \geq 0$ to be equivalent to:
\begin{align*}
    \{ c_l & := a_j == b_j, \\
    c_{l+1} & := a_{j+1} == b_{j+1}, \\
    ..., \\
    c_{l+k-j} & := a_{k} == b_{k} \} \\
\end{align*}
In particular, $c$ is a tuple of boolean equations with cardinality $|c| = k-j$,
defining an equality relation between input sets $a$ and $b$, at index starting
at $j$ and ending at $k$ (inclusive).

@@

\subsubsection{Constraint-Based}
\begin{itemize}
    \item Simple encoding
    \item Advanced constraints
    \item Time difference
\end{itemize}
testing
\subsubsection{Differential-Independent}
\subsubsection{State-Independent}
h1: f7f3fe7437e33bee3590392936aeab37e0b848329173bf0f08a252a0281e3284f4e4fd90626c4cbb3350ac3e36694d34703d31498806e520608c6738da70170b:40a8b995bd89b97662b4516e1df317f2
h2: f7f3fe7437e33b6e3590399936aeab37e0b848329173bf0f08a252a0281e3284f4e4fd90626c4cbb3350ac3e36694d34703d30498806e520608c6738da70170b:40a8b995bd89b97662b4516e1df317f2
h3: f7f3fe7437e33bee3590392936aeab37e0b848329173bf0f08a252a0281e3284f4e4fd90626c4cbb3350ac3e36694d34703d31498806e520608c6738da70170b:aabcc2e21ea0e529178ca7dcb21240a2
h4: f7f3fe7437e33b6e3590399936aeab37e0b848329173bf0f08a252a0281e3284f4e4fd90626c4cbb3350ac3e36694d34703d30498806e520608c6738da70170b:aabcc2e21ea0e529178ca7dcb21240a2
c Total time               : 127.12

\subsubsection{Malleable}
\begin{itemize}
    \item ASCII
    \item JSON
    \item substring
\end{itemize}
Testing
\subsubsection{Additional Notes}

@@

\subsubsection{State Independent}
h1: 13d340f35ff2c0ace19bcd73f5c153aa816105701b5050a2c7e0ab260adf14026e0a20b1f39a8f4be37e60aad37d5b1100140082c228676c4d1c8098c095d023:746094a6c1cf11f96614c893f6b54d67
h2: 13d340035ff2c0ace19bcdf3f5c153aa816105f01b5050a2c7e0ab260adf14026e0a2031f39a8f4be37e60aad37d5b1100140002c228676c4d1c8098c095d023:746094a6c1cf11f96614c893f6b54d67
h3: 13d340f35ff2c0ace19bcd73f5c153aa816105701b5050a2c7e0ab260adf14026e0a20b1f39a8f4be37e60aad37d5b1100140082c228676c4d1c8098c095d023:ef64866bf9eedb5a723d7153bcce781f
h4: 13d340035ff2c0ace19bcdf3f5c153aa816105f01b5050a2c7e0ab260adf14026e0a2031f39a8f4be37e60aad37d5b1100140002c228676c4d1c8098c095d023:ef64866bf9eedb5a723d7153bcce781f
c Total time               : 165.32

\section{MD5}
\subsection{Brief History and Overview}
\subsection{Wang's MD5 Collisions}
\subsection{Stevens's MD5 Collisions}
\subsection{Xie's MD5 Collisions}
\subsection{Sasaki's MD5 Collisions}
